### 2019-07-10 0/exams/m1-practice.tex
Start of m1 practice.
Practice midterm 1.

\begin{center}
  {\Large {\bf CMSC 131A, Midterm 1}\\
    \censor{{\bf SOLUTION}} \ \\
    \ \\
    Fall 2018 \ \\
  }
\end{center}

@@

\noindent
UID: \verb|__________________________________________|

\begin{center}
\begin{tabular}{| c | c |}
  \hline
  Question & Points \\ \hline \hline
  1 & 10 \\ \hline
  2 & 10 \\ \hline
  3 & 15 \\ \hline
  4 & 15 \\ \hline
  5 & 40 \\ \hline
  Total: & 90\\
  \hline
\end{tabular}
\end{center}

\noindent
This test is open-book, open-notes, but you may not use any computing
device other than your brain and may not communicate with anyone.
You have 50 minutes to complete the test.

@@

(define (swap s i)
  (string-append (substring s i)
                 (substring s 0 i)))
\end{SaveVerbatim}

@@

\newpage
\noindent
{\bf Problem 4 (15 points).}  
%
Design a program that takes a list of numbers and produces the sum of
positive numbers in the list.  For example, if the list contains 3, 4,
-1, 0, -2, and 5, the sum of positive numbers (3+4+5) is 12.  You may
use the following data definition:

\begin{verbatim}
;; A LoN (list of numbers) is one of:
;; - '()
;; - (cons Number LoN)
;; Interp: an arbitrarily long list of numbers
\end{verbatim}

@@

;; sum-pos : LoN -> Number
;; Count the number of positive numbers in list
(check-expect (sum-pos '()) 0)
(check-expect (sum-pos (list 1 -2 8 0)) 9)
(define (sum-pos lon)
  (cond [(empty? lon) 0]
        [(cons? lon)
         (if (< (first lon) 0)
             (sum-pos (rest lon))
             (+ (first lon) (sum-pos (rest lon))))]))
\end{SaveVerbatim}

@@

\noindent
{\bf Problem 5 (40 points).} You've been hired by Nintendo to design
their next blockbuster game: Dots.  It consists of, well, dots: lots
and lots of dots, each of which occupies a coordinate in space.  You
come up with the following data definitions for representing a single
dot and a collection of dots:

\begin{verbatim}
;; A Dots is one of:
;; - empty
;; - (cons Dot Dots)
;; Interp: collection of an arbitrary number of dots

;; A Dot is a (make-posn Number Number)
;; Interp: a dot's position in space
\end{verbatim}

\vspace{1em}

\noindent
{\bf Problem 5(a) [10 points].} Write templates for \verb|Dot| and
\verb|Dots| functions.

@@

(define (dot-template d)
  (... (posn-x d)
       (posn-y d)
       ...))

(define (dots-template ds)
  (cond [(empty? ds) ...]
        [(cons? ds) (... (dot-template (first ds))
                         (dots-template (rest ds)) ...)]))
\end{SaveVerbatim}

\vskip 18em

\censor{\vspace{-15em}}
\censor{%
\noindent\BUseVerbatim{VerbEnv}}

\noindent
{\bf Problem 5(b) [10 points].} Write a stub for the following function:

\begin{verbatim}
;; nw : Dots -> Dots
;; Move given dots in the northwest direction one unit
(check-expect (nw empty) empty)
(check-expect (nw (cons (make-posn 3 4) empty))
              (cons (make-posn 4 5) empty))
\end{verbatim}

(define (nw ds) empty)

\end{SaveVerbatim}

\censor{%
\noindent\BUseVerbatim{VerbEnv}}

\noindent
{\bf Problem 5(c) [20 points].}
%
Design two functions: {\tt dot-flip} reflects a given dot over the
y-axis, for example, $(3,4)$ becomes $(-3,4)$, and {\tt dots-flip}
reflects a collection of dots over the y-axis.
\begin{SaveVerbatim}{VerbEnv}

;; SOLUTION:

;; dot-flip : Dot -> Dot
;; Flip dot over y-axis
(check-expect (dot-flip (make-posn 3 4)) (make-posn -3 4))
(define (dot-flip d)
  (make-posn (- (posn-x d)) (posn-y d)))

;; dots-flip : Dots -> Dots
;; Flip dots over y-axis
(check-expect (dots-flip empty) empty)
(check-expect (dots-flip (cons (make-posn 3 4) empty))
              (cons (make-posn -3 4) empty))
(define (dots-flip ds)
  (cond [(empty? ds) empty]
        [(cons? ds) 
         (cons (dot-flip (first ds))
               (dots-flip (rest ds)))]))                   
\end{SaveVerbatim}

@@

\noindent
[Extra space, should you need it.]

%% Design a program that consumes an arbitrarily large collection of
%% numbers and produces the count of non-negative numbers in the
%% collection.

%% \begin{SaveVerbatim}{VerbEnv}

%% SOLUTION:

%% ;; A LoN (list of numbers) is one of:
%% ;; - '()
%% ;; - (cons Number LoN)

%% ;; count-non-neg : LoN -> Natural
%% ;; Count the number of non-negative numbers in the given list.
%% (check-expect (count-non-neg '()) 0)
%% (check-expect (count-non-neg (cons -1 (cons 1 (cons -2 '())))) 1)
%% (define (count-non-neg lon)
%%   (cond [(empty? lon) 0]
%%         [(cons? lon) 
%%          (+ (if (<= 0 (first lon)) 1 0)
%%             (count-non-neg (rest lon)))]))
%% \end{SaveVerbatim}

%% \censor{%
%% \noindent
%% \BUseVerbatim{VerbEnv}}

>>>>

\begin{center}
  {\Large {\bf CMSC 198Q, Midterm 1 (PRACTICE)}\\
    \censor{{\bf SOLUTION}} \ \\
    \ \\
    Summer 2019 \ \\
  }
\end{center}

@@

\begin{center}
\begin{tabular}{| c | c |}
  \hline
  Question & Points \\ \hline \hline
  1 & 10 \\ \hline
  2 & 10 \\ \hline
  3 & 15 \\ \hline
  4 & 10 \\ \hline
  5 & 20 \\ \hline
  Total: & 65\\
  \hline
\end{tabular}
\end{center}

\noindent
This test is open-book, open-notes, but you may not use any computing
device other than your brain and may not communicate with anyone.
You have 60 minutes to complete the test.

@@

(define (dear p)
  (string-append "Dear " 
                 (title-abbrv (person-title p)) 
                 " " 
                 (person-last p) 
                 ":"))

;; title-string : Title -> String
;; Render title as a string abbreviation
(check-expect (title-abbrv "d") "Dr.")
(check-expect (title-abbrv "r") "Mr.")
(check-expect (title-abbrv "s") "Ms.")
(define (title-abbrv t)
  (cond [(string=? t "d") "Dr."]
        [(string=? t "r") "Mr."]
        [(string=? t "s") "Ms."]))
\end{SaveVerbatim}

@@

\newpage
\noindent
{\bf Problem 4 (10 points).}
%
Do you ever mispell words like ``peice'' or ``acheive'' (which are
correctly spelled ``piece'' and ``achieve'')?  Turns out a lot of
people make \emph{transposition} errors like this when typing.  Based
on this insight, you decide to make a next generation messaging app
that helps users by giving them a \emph{transpose} operation.  The
idea is that as a user types, they can place their cursor between two
letters that need to be transposed and invoke {\tt transpose}.  To
implement this feature, you define the following function for
transposing letters at a given position in a string that has at
least 2 letters in it:
\begin{verbatim}
;; transpose : String Index -> String
;; Transpose characters at left and right of index i.
;; Assumes string has length >= 2 and 1<=i, i+1<=length.
(check-expect (transpose "ab" 1) "ba")
(check-expect (transpose "peice" 2) "piece")
(check-expect (transpose "student's" 8) "students'")
(define (transpose s i) ...)
\end{verbatim}

\noindent
Give a correct definition for \verb|transpose|.  (You only need to
provide code, not the design steps.)

@@

(define (transpose s i)
  (string-append (substring s 0 (sub1 i))
                 (substring s i (add1 i))
                 (substring s (sub1 i) i)
                 (substring s (add1 i))))
\end{SaveVerbatim}

@@

\newpage
\noindent
{\bf Problem 5 (20 points).}  Growing tired of Bug, you
decide to build a new video game.  Part of the game consists of flying
billiard balls that move in straight lines at varying velocities until
hitting other balls or bouncing off walls and other obstacles.  After
making some sketches, you decide on the following data representation
for billiard balls:

\begin{verbatim}
;; A BB is a:
;; (make-bb Color 
;;          (make-posn Integer Integer) 
;;          (make-vel Integer Integer))

;; A Color is one of:
;; - "red"
;; - "yellow"
;; - "green"

(define-struct bb (color center vel))
(define-struct vel (deltax deltay))
\end{verbatim}
The interpretation of a \verb|BB| is that the color is the color of
the ball, the \verb|posn| is the location of the center of the ball,
and the \verb|vel| structure describes the ball's velocity as a change
along the $x$-axis (\verb|deltax|) and $y$-axis (\verb|deltay|) in one
clock tick.

\noindent
Design a function called \verb|tock : BB -> BB| that calculates where
a given billiard ball will be, based on its velocity, after one tick
of the clock and assuming it does not encounter any obstacle.

@@

;; tock : BB -> BB
;; Move ball one tick based on velocity
(check-expect (tock (make-bb "red" (make-posn 2 5) (make-vel 1 -2)))
              (make-bb "red" (make-posn 3 3) (make-vel 1 -2)))
(define (tock bb)
  (make-bb (bb-color bb)
           (make-posn (+ (posn-x (bb-center bb))
                         (vel-deltax (bb-vel bb)))
                      (+ (posn-y (bb-center bb))
                         (vel-deltay (bb-vel bb))))
           (bb-vel bb)))
\end{SaveVerbatim}

@@
